Algomusicology, ????, Profit

### 2016-05-29 paper.tex
Tie the different parts together more explicitly.

This paper will argue that it can. It will do so by citing examples of existing art works that are algorithmic in nature, and
also by using algomusicological concepts to create a work that challenges our present conception of musical authorship and
copyright.

@@

@@

This results in something very similar to the Library of Babel as imagined by Jorge Luis Borges\cite{The Library of Babel}. In his
vast fictional library, thousands of hexagonal rooms house all possible books of four hundred and ten pages. Melancholic
librarians spend their entire lives toiling in its maddiningly symetrical depths.

@@

Borges appreciated this paradox. His anonymous narrator reports that on ``some shelf in some hexagon (men reasoned)
there must exist a book which is the formula and perfect compendium of all the rest". In other words, that one book
epitomises the complexity of the entire library, though it is part of the library.

Any of the many books that contain our \verb|library-of-babel| function match the narrator's description,
though the books that also contain the rest of this paper must surely provide more helpful context.

@@

In Carl Sagan's novel `Contact'\cite{Contact}, he imagines that a message to humanity is encoded in the digits of pi.
If we design an appropriate encoding scheme, we could do a similar thing and treat the digits of pi as music. However,
we would encounter a similar problem to Borges' librarians - we suspect that every sequence of data we could possibly want
is in there somewhere, but we don't know where to look.

@@

The Criminal Justice and Public Order Act 1994 was an attempt by the UK Government to curb rave culture. The musicological
dilemma encountered by the government of the day was how to define raves. They settled on prohibiting gatherings where
``a succession of repetitive beats" were played\cite{Joseph Gallivan on pop}.

@@

@@

As an artform, music is an ideal subject for complexity analysis. Other media can also engage with complexity theory, but
few are as tractable to complexity theory's techniques. Sagan and Borges may have written fiction that explores the implications
of algorithmic creation, but it would be difficult to analyse 'Contact' and 'The Library of Babel' through Kolmogorov's lens.

@@

\bibitem{Overtone} Aaron, Sam and Rose, Jeff: Overtone (\url{http://overtone.github.io})
\bibitem{Anti EP} Autechre: Anti EP
\bibitem{Air on the G String} Wilhelmj, August arranging Bach, Johann Sebastian: Air on the G String
\bibitem{The Library of Babel} Borges, Jorge Luis: The Library of Babel
\bibitem{The construction of decimals normal in the scale of ten} Champernowne, David: The construction of decimals normal in the scale of ten
\bibitem{Gnomes} Brady, Pam, Parker, Trey and Stone, Matt. Episode 17, season 2 of South Park
\bibitem{Got to Give It Up} Got to Give It Up: Gaye, Marvin
\bibitem{Leipzig} Ford, Chris: Leipzig
\bibitem{Joseph Gallivan on pop} Gallivan, Joseph: Joseph Gallivan on pop (\url{http://www.independent.co.uk/life-style/jospeh-gallivan-on-pop-1414727.html})
\bibitem{Clojure} Hickey, Rich: Clojure
\bibitem{On Tables of Random Numbers}  Kolmogorov, Andrey: On Tables of Random Numbers
\bibitem{Midje} Marick, Brian: Midje
\bibitem{Analysis by Compression} Meredith, David: Analysis by Compression
\bibitem{USA Today} Oldenburg, Ann: 'Blurred Lines' jury finds for Marvin Gaye (\url{http://www.usatoday.com/story/life/music/2015/03/10/blurred-lines-trial-verdict/24492431/})
\bibitem{Contact} Sagan, Carl: Contact
\bibitem{Blurred Lines} Williams, Pharrel and Thicke, Robin: Blurred Lines

>>>>

This paper will argue that it can. It will do so by reviewing existing art works that are algorithmic in nature, and
also by using algomusicological concepts to create a new work that challenges our present conception of musical authorship and
copyright. As befits its subject matter, this paper will incorporate conventions and style from literary essays and musicology
papers.

@@

Music is widely regarded to have mathematical properties, but literature also contains examples of works where artists have engaged
with algorithmic and information theoretical concepts. One striking case is the Library of Babel as imagined by Jorge Luis
Borges\cite{The Library of Babel}. We will examine the story in a moment, but first we must define the mathematical concept
that it revolves around - the Kleene star.

@@

This results in something very similar to Borges' Library of Babel. In this vast fictional library, thousands of hexagonal
rooms house all possible books of four hundred and ten pages. Melancholic librarians spend their entire lives toiling in its
maddiningly symetrical depths.

@@

Borges appreciated this paradox. His anonymous narrator reports that on ``some shelf in some hexagon (men reasoned)
there must exist a book which is the formula and perfect compendium of all the rest". In other words, that one book
epitomises the complexity of the entire library, though it is part of the library. Any of the many books that contain our
\verb|library-of-babel| function match the narrator's description, though the books that also contain the rest of this paper
must surely provide more helpful context.

Through his Library of Babel, Borges is playing with our notions of invention versus discovery. Mathematicians and computer
scientists struggle with whether they are the inventors of new theories or discoverers. Philip Wadler, for example,
argues that the lambda calculus is such a fundamental concept that it deserves to be considered a universal concept that
was discovered\cite{Propositions as Types}. By positing a library where any work may be found if you know where to look, Borges
suggests that even literary authorship could be considered a search problem and therefore a matter of discovery rather than
creation.

@@

In Carl Sagan's novel `Contact'\cite{Contact}, he takes the concept of discovering meaning in naturally occurring sources of
information even further than Borges. Firstly, Sagan imagines that SETI detects a transmission consisting of a series of prime
numbers from the the Vega star system. This prompts an exploration of how to distinguish between a signal created by an
intelligence and one arising from natural symmetry.

Later a message to humanity is discovered encoded in the digits of pi, placed there by beings who are able to manipulate the
structure of the universe itself. This development questions the dichotomy between invented and discovered information altogether,
by imagining a mathematical quantity as a medium for meaningful communication.

If we design an appropriate encoding scheme, we could do a similar thing and interpret the digits of pi as music. However,
we would encounter a similar problem to Borges' librarians - we suspect that every sequence of data we could possibly want
is in there somewhere, but we don't know where to look.

@@

Complexity and information theory have been used to speculate on grand questions like the place of humanity in the universe,
but they have also been used by artists to comment on specific social issues. The Criminal Justice and Public Order Act 1994
was an attempt by the UK Government to curb rave culture. The musicological dilemma encountered by the government of the day
was how to define raves. They settled on prohibiting gatherings where ``a succession of repetitive beats" were
played\cite{Joseph Gallivan on pop}.

@@

We will now take the ideas explored in previous sections and rework them into a novel contribution to the politics of
copyright and creativity.

@@

As an artform, music is an ideal subject for complexity analysis. Other media can also engage with complexity theory, but
few are as tractable to complexity theory's techniques. Sagan and Borges may have written fiction that explores the implications
of algorithmic creation, but it would be difficult to analyse 'Contact' and 'The Library of Babel' themselves through
Kolmogorov's lens.

Algomusicology is part of a rich tradition of information theoretical art and criticism. Music's deep structure and cultural
significance give algomusicology great potential as a source of artistic analysis and inspiration.

@@

\bibitem{Overtone} Aaron, Sam and Rose, Jeff: Overtone (\url{http://overtone.github.io})
\bibitem{Anti EP} Autechre: Anti EP
\bibitem{Air on the G String} Wilhelmj, August arranging Bach, Johann Sebastian: Air on the G String
\bibitem{The Library of Babel} Borges, Jorge Luis: The Library of Babel
\bibitem{The construction of decimals normal in the scale of ten} Champernowne, David: The construction of decimals normal in the scale of ten
\bibitem{Gnomes} Brady, Pam, Parker, Trey and Stone, Matt. Episode 17, season 2 of South Park
\bibitem{Got to Give It Up} Got to Give It Up: Gaye, Marvin
\bibitem{Leipzig} Ford, Chris: Leipzig
\bibitem{Joseph Gallivan on pop} Gallivan, Joseph: Joseph Gallivan on pop (\url{http://www.independent.co.uk/life-style/jospeh-gallivan-on-pop-1414727.html})
\bibitem{Clojure} Hickey, Rich: Clojure
\bibitem{On Tables of Random Numbers}  Kolmogorov, Andrey: On Tables of Random Numbers
\bibitem{Midje} Marick, Brian: Midje
\bibitem{Analysis by Compression} Meredith, David: Analysis by Compression
\bibitem{USA Today} Oldenburg, Ann: 'Blurred Lines' jury finds for Marvin Gaye (\url{http://www.usatoday.com/story/life/music/2015/03/10/blurred-lines-trial-verdict/24492431/})
\bibitem{Contact} Sagan, Carl: Contact
\bibitem{Propositions as Types} Wadler, Philip.
\bibitem{Blurred Lines} Williams, Pharrel and Thicke, Robin: Blurred Lines
